---
title: "Mites distribution in apple orchard, and the effect of predatory mites, as biological control"
output:
  html_document:
    code_folding: hide
    theme: cerulean
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
      df_print: paged
editor_options: 
  chunk_output_type: console
---

```{=html}
<style type="text/css">
  body{
  font-size: 12pt;
}
</style>
```
## load libraries
```{r libraries, echo=TRUE, message=FALSE, warning=FALSE}
library("tidyverse")
library("plyr")
library("dplyr")
library("ggplot2")
#library("scales")
#library("ggpubr")
#library("gridExtra")
#library("grid")
#library("GGally")
library("data.table")
library("stringr")
library("janitor")
library("knitr")
library("kableExtra")
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
                 #    fig.width = 20,
                  #    fig.asp = 0.6,
                   #   out.width = "100%")
```

The aim of the study is to look at the distribution of herbivore mites:  
- [*Tetranychus sp.*](https://www.biobee.com/pests/two-spotted-spider-mite/)  (Yellow/red mite)   
- [*Panonychus ulmi*](http://www.agri.huji.ac.il/mepests/pest/Panonychus_ulmi/)  (European red mite)   
- [*Bryobia rubrioculus*](http://www.agri.huji.ac.il/mepests/pest/Bryobia_rubrioculus/) (brwon mite)  

and the predatory mites:  
- [*Typhlodromus athiasae*](http://www.agri.huji.ac.il/mepests/enemy/Typhlodromus_athiasae/)  (local species)   
- [*Neoseiulus californicus*](https://entnemdept.ufl.edu/creatures/beneficial/neoseiulus_californicus.htm) (artificially introduced 30 years ago, and augmented in the current study).  

the predatory mites, *Neoseiulus californicus*, artificially reared at Biobee Sde Eliyahu, were applied on ___ via the sachets method, in 3 concentrations:  

- **High**: one sachet per tree, 8 rows.    
- **Low**: one sachet every 5 trees, 7 rows.    
- **Control**:  no sachets applied, 7 rows.     

---------

**The study design was as follow:**  
A total of ~9 dunams, containing 22 rows, in each row one type of cultivar: Gala, Gold, Pink or Grani Smith.  

In each row, 68 trees, except for Smith, which is planted in higher gaps and so 48 trees are planted per row.   


```{r}
map <- read.csv(check.names=FALSE, "/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/data/study_map.csv")
kable(map, caption = "Study design, each column is one row") %>%
  column_spec(1:23,border_left = T, border_right = T) %>%
kable_styling()
```

the mites population was monitored every ~2 weeks, by inspecting 10 leaves under a magnified glass, and observing adult and eggs (5 outer and 5 inner leaves).   

the level of infestation was classified as follow:  
- **"0"**: no mites observed  
- **"1-5"**  
- **"5-10"**  
- **10+"**, more then 10 mites/eggs per leave  

5-8 trees were monitored per row, and the trees were marked and re-inspected each time.  

---------
*The study was conducted in collaboration of [Shamir Research Institute](http://www.sri.org.il/Home), and [Biobee company](https://www.biobee.com/).*   
--------
### load data
```{r}
meta <- read.csv("/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/data/18jun.csv")
design <- read.csv("/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/data/design.csv")

data <- left_join(design,meta, by = "row") %>% # match the cultivar name to each row
  filter(row !="23") %>% # remove row 23 (we checked only once)
  mutate(across(everything(), as.character)) 

 data[data == ""] <- NA
 data[is.na(data)] <- 0

 #write the arranged data as csv file
#write_csv(x= data, file="/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/outfiles/data.csv")

#data <- left_join(design,meta, by = "row") %>% # match the cultivar name to each row
 # filter(row !="23")
 #data <- data %>% mutate(across(-1,  ~ as.numeric(replace(., . == '', 0))))  %>% 
#  mutate(across(everything(), as.character)) %>%
#replace(is.na(.),0)
```

## Adult mites distribution on leaves
### on outer leaves
```{r}
# count adult mites 
#on 5 outer leaves:
out_leaves <- data %>% 
  dplyr::select(-contains('egg')) %>%
  dplyr::select(-starts_with('inner')) %>%
  pivot_longer(cols = starts_with('outer')) %>% 
  separate(name, c('leave_position', 'leave_number', 'mite_species'))  %>% 
  group_by( mite_species, tree, strain, treatment) %>%
  dplyr::count(value, .drop = FALSE)

# plot the distribution 
out_leaves$value <- factor(out_leaves$value, levels = c("0","1-5","5-10","10+"))

out_leaves %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "mite infestation distribution on outer leaves,
comparison between the 4 species") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + scale_fill_manual(values=c("#66b032", "#FFC300","#FF5733","#C70039")) +
  facet_wrap(~mite_species)

out_leaves %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "mite infestation distribution on outer leaves
         comparison between the 4 apples cultivars") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + scale_fill_manual(values=c("#66b032", "#FFC300","#FF5733","#C70039")) +
  facet_wrap(~strain)
```

### on Inner leaves
```{r}
# count adult mites 
#on 5 outer leaves:
in_leaves <- data %>% 
  dplyr::select(-contains('egg')) %>%
  dplyr::select(-starts_with('outer')) %>%
  pivot_longer(cols = starts_with('inner')) %>% 
  separate(name, c('leave_position', 'leave_number', 'mite_species'))  %>% 
  group_by( mite_species, tree, strain, treatment) %>%
  dplyr::count(value, .drop = FALSE)

# plot the distribution 
in_leaves$value <- factor(in_leaves$value, levels = c("0","1-5","5-10","10+"))

in_leaves %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "mite infestation distribution on inner leaves
         comparison between the 4 mite species") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + scale_fill_manual(values=c("#66b032", "#FFC300","#FF5733","#C70039")) +
  facet_wrap(~mite_species)

in_leaves %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "mite infestation distribution on inner leaves
         comparison between the 4 apple cultivars") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + scale_fill_manual(values=c("#66b032", "#FFC300","#FF5733","#C70039")) +
  facet_wrap(~strain)
```

## Mites' eggs distribution on leaves
### on outer leaves
```{r}
# count  mites eggs on 5 outer leaves:
out_leaves_eggs <- data %>% 
  dplyr::select( c(row, treatment, strain, tree),contains('Egg')) %>%
  dplyr::select(-starts_with('inner')) 

# make a list of the columns to keep:
keep <- out_leaves_eggs %>% dplyr::select(starts_with('outer')) %>%
  colnames()
out_leaves_eggs <- out_leaves_eggs %>% 
  mutate_at(keep, as.character) %>%
  pivot_longer(cols = starts_with('outer')) %>% 
  separate(name, c('leave_position', 'leave_number', 'egg_type'))  %>%
  group_by(egg_type, tree, strain, treatment) %>%
  dplyr::count(value, .drop = FALSE)

# plot the distribution 
out_leaves_eggs$value <- factor(out_leaves_eggs$value, levels = c("0","1-5","5-10","10+"))

out_leaves_eggs %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "Egg distribution on outer leaves") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + scale_fill_manual(values=c("#66b032", "#FFC300","#FF5733","#C70039")) +
  facet_wrap(~egg_type)

out_leaves_eggs %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "Egg distribution on outer leaves") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + scale_fill_manual(values=c("#66b032", "#FFC300","#FF5733","#C70039")) +
  facet_wrap(~strain)
```

### on inner leaves
```{r}
# count  mites eggs  
#on 5 inner leaves:
in_leaves_eggs <- data %>% 
  dplyr::select( c(row, treatment, strain, tree),contains('egg')) %>%
  dplyr::select(-starts_with('outer')) %>%
  pivot_longer(cols = starts_with('inner')) %>% 
  separate(name, c('leave_position', 'leave_number', 'egg_type'))  %>% 
  group_by(egg_type, tree, strain, treatment) %>%
  dplyr::count(value, .drop = FALSE)

# plot the distribution 
in_leaves_eggs$value <- factor(in_leaves_eggs$value, levels = c("0","1-5","5-10","10+"))

in_leaves_eggs %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "Egg distribution on outer leaves") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + scale_fill_manual(values=c("#66b032", "#FFC300","#FF5733","#C70039")) +
  facet_wrap(~egg_type)

in_leaves_eggs %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "Egg distribution on outer leaves") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + scale_fill_manual(values=c("#66b032", "#FFC300","#FF5733","#C70039")) +
  facet_wrap(~strain)
```
