---
title: "Mites distribution in apple orchard, and the effect of predatory mites, as biological control"
output:
  html_document:
    code_folding: hide
    theme: cerulean
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
      df_print: paged
editor_options: 
  chunk_output_type: console
---

```{=html}
<style type="text/css">
  body{
  font-size: 12pt;
}
</style>
```
## load libraries
```{r libraries, echo=TRUE, message=FALSE, warning=FALSE}
library("tidyverse")
library("plyr")
library("dplyr")
library("ggplot2")
library("RColorBrewer")
#library("scales")
#library("ggpubr")
#library("gridExtra")
#library("grid")
#library("GGally")
library("data.table")
library("stringr")
library("janitor")
library("knitr")
library("kableExtra")
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
                 #    fig.width = 20,
                  #    fig.asp = 0.6,
                   #   out.width = "100%")
```

The aim of the study is to look at the distribution of herbivore mites:  
- [*Tetranychus sp.*](https://www.biobee.com/pests/two-spotted-spider-mite/)  (Yellow/red mite)   
- [*Panonychus ulmi*](http://www.agri.huji.ac.il/mepests/pest/Panonychus_ulmi/)  (European red mite)   
- [*Bryobia rubrioculus*](http://www.agri.huji.ac.il/mepests/pest/Bryobia_rubrioculus/) (brwon mite)  

and the predatory mites:  
- [*Typhlodromus athiasae*](http://www.agri.huji.ac.il/mepests/enemy/Typhlodromus_athiasae/)  (local species)   
- [*Neoseiulus californicus*](https://entnemdept.ufl.edu/creatures/beneficial/neoseiulus_californicus.htm) (artificially introduced 30 years ago, and augmented in the current study).  

the predatory mites, *Neoseiulus californicus*, artificially reared at Biobee Sde Eliyahu, were applied on ___ via the sachets method, in 3 concentrations:  

- **High**: one sachet per tree, 8 rows.    
- **Low**: one sachet every 5 trees, 7 rows.    
- **Control**:  no sachets applied, 7 rows.     

---------

**The study design was as follow:**  
A total of ~9 dunams, containing 22 rows, in each row one type of cultivar: Gala, Gold, Pink or Grani Smith.  

In each row, 68 trees, except for Smith, which is planted in higher gaps and so 48 trees are planted per row.   


```{r}
map <- read.csv(check.names=FALSE, "/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/data/study_map.csv")
kable(map, caption = "Study design, each column is one row") %>%
  column_spec(1:23,border_left = T, border_right = T) %>%
kable_styling()
```

the mites population was monitored every ~2 weeks, by inspecting 10 leaves under a magnified glass, and observing adult and eggs (5 outer and 5 inner leaves).   

the level of infestation was classified as follow:  
- **"0"**: no mites observed  
- **"1-5"**  
- **"5-10"**  
- **10+"**, more then 10 mites/eggs per leave  

5-8 trees were monitored per row, and the trees were marked and re-inspected each time.  

---------
 * The study was conducted in collaboration of [Shamir Research Institute](http://www.sri.org.il/Home), and [Biobee company](https://www.biobee.com/).*   
--------

/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/data/5_jun_num.csv
## 5 june 2023
#### as numeric
```{r}
meta_5jun_num <- read.csv("/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/data/5_jun_num.csv")
design <- read.csv("/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/data/design.csv")

data_5jun_num  <- right_join(design,meta_5jun_num, by = "row") %>% 
  # match the cultivar name to each row , and remove rows 21,22,23 (not tested in this date)
  mutate(across(everything(), as.character)) 

 data_5jun_num[data_5jun_num == ""] <- NA
 data_5jun_num[is.na(data_5jun_num)] <- 0
 
all_leaves_5jun<- data_5jun_num %>% 
  pivot_longer(cols = contains('_')) %>% 
  separate(name, c('mite_species','leave_number'))
 
# order the levels 
all_leaves_5jun$mite_species <- factor(all_leaves_5jun$mite_species, level=c("tetranychus", "panonychus", "bryobia", "predatory"))
all_leaves_5jun$treatment <- factor(all_leaves_5jun$treatment, level=c("control", "low", "high"))

all_leaves_5jun %>% 
  group_by(treatment, mite_species) %>%
  dplyr::summarise(mean=mean(value),
            sd=sd(value), .groups = 'drop') %>%
  ggplot(aes(y=mean, x=treatment, fill = treatment)) + 
 geom_bar(position = "dodge",
           stat = "identity") +
        labs(title = "Mite count per leaf, 5/6/2023") +
  #geom_errorbar(aes(ymin=0, ymax=mean+sd)) +
  facet_wrap(~mite_species, scales ="free") + 
  theme_classic() 


# by apple strain 
 all_leaves_5jun %>% 
  group_by(treatment, mite_species,strain) %>%
  dplyr::summarise(mean=mean(value),
            sd=sd(value), .groups = 'drop') %>%
  ggplot(aes(y=mean, x=treatment, fill = strain)) + 
 geom_bar(position = "dodge",
           stat = "identity")+
    # geom_errorbar(position = "dodge",aes(ymin=0, ymax=mean+sd)) +
    labs(title = "Mite count per leaf, 5/6/2023") +
  facet_wrap(~mite_species, scales ="free") + 
  theme_classic() 

#in boxplot the outliers really bias the plots, making it hard to detect differences
# all_leaves_5jun %>% 
#  group_by(treatment, mite_species) %>%
#  ggplot(aes(y=value, x=treatment, fill = treatment)) + 
#   geom_boxplot() +
#  facet_wrap(~mite_species) + 
#  theme_classic() 
```
 
##### spatial distribution - bubble plot  
```{r}
df = all_leaves_5jun %>%
  mutate(tree = as.numeric(tree)) %>%
    mutate(row = as.numeric(row))

 
df %>%  dplyr::filter(mite_species ==c ("tetranychus", "predatory")) %>%
 arrange(desc(value)) %>%
ggplot(aes(x=row, y=tree, size=value, color=mite_species)) +
    geom_point(alpha=0.5) +
    scale_size(range = c(2, 20), name="Mite count") +
    scale_y_continuous(breaks=seq(0,70,5)) +   
     scale_x_reverse(breaks=seq(20,1)) +
   theme(text = element_text(size=20)) +
    labs(title = "Mite count per leaf") +
   theme(panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey90"),  panel.grid.minor.x = element_blank(),  panel.grid.minor.y = element_blank()) 
```
 
## 18 June 2023
#### as numeric
```{r}
meta_18jun_num <- read.csv("/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/data/18_jun_num.csv")
design <- read.csv("/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/data/design.csv")

data_18jun_num <- right_join(design,meta_18jun_num, by = "row") %>% 
  mutate(across(everything(), as.character)) 

 data_18jun_num[data_18jun_num == ""] <- NA
 data_18jun_num[is.na(data_18jun_num)] <- 0

# count adult mites 
#on 5 outer leaves:
out_leaves <- data_18jun_num %>% 
   #dplyr::mutate(across(everything(), as.character)) %>%
  dplyr::select(-contains('egg')) %>%
  dplyr::select(-starts_with('inner')) %>%
  pivot_longer(cols = starts_with('outer')) %>% 
  separate(name, c('leave_position', 'leave_number', 'mite_species')) 

in_leaves <- data_18jun_num %>% 
   #dplyr::mutate(across(everything(), as.character)) %>%
  dplyr::select(-contains('egg')) %>%
  dplyr::select(-starts_with('outer')) %>%
  pivot_longer(cols = starts_with('inner')) %>% 
  separate(name, c('leave_position', 'leave_number', 'mite_species')) 

all_leaves_18jun = bind_rows(in_leaves, out_leaves)

# order the levels 
all_leaves_18jun$mite_species <- factor(all_leaves_18jun$mite_species, level=c("tetranychus", "panonychus", "bryobia", "predatory"))
all_leaves_18jun$treatment <- factor(all_leaves_18jun$treatment, level=c("control", "low", "high"))

all_leaves_18jun %>% 
  group_by(treatment, mite_species) %>%
  dplyr::summarise(mean=mean(value),
            sd=sd(value), .groups = 'drop') %>%
  ggplot(aes(y=mean, x=treatment, fill = treatment)) + 
 geom_bar(position = "dodge",
           stat = "identity") +
        labs(title = "Mite count per leaf, 18/6/2023") +
  #geom_errorbar(aes(ymin=0, ymax=mean+sd)) +
  facet_wrap(~mite_species, scales ="free") + 
  theme_classic() 
  
# by leave position
 all_leaves_18jun %>% 
  group_by(treatment, mite_species,leave_position) %>%
  dplyr::summarise(mean=mean(value),
            sd=sd(value), .groups = 'drop') %>%
  ggplot(aes(y=mean, x=treatment, fill = leave_position)) + 
 geom_bar(position = "dodge",
           stat = "identity")+
  #    geom_errorbar(position = "dodge",aes(ymin=0, ymax=mean+sd)) +
    labs(title = "Mite count per leaf, 18/6/2023") +
  facet_wrap(~mite_species, scales ="free") + 
  theme_classic() 

# by apple strain 
 all_leaves_18jun %>% 
  group_by(treatment, mite_species,leave_position,strain) %>%
  dplyr::summarise(mean=mean(value),
            sd=sd(value), .groups = 'drop') %>%
  ggplot(aes(y=mean, x=treatment, fill = strain)) + 
 geom_bar(position = "dodge",
           stat = "identity")+
    # geom_errorbar(position = "dodge",aes(ymin=0, ymax=mean+sd)) +
    labs(title = "Mite count per leaf, 18/6/2023") +
  facet_wrap(~mite_species, scales ="free") + 
  theme_classic() 

#in boxplot the outliers really bias the plots, making it hard to detect differences
#  all_leaves_18jun %>% 
#  group_by(treatment, mite_species) %>%
#  ggplot(aes(y=value, x=treatment, fill = treatment)) + 
#   geom_boxplot() +
#  facet_wrap(~mite_species) + 
#  theme_classic() 
```

##### spatial distribution - bubble plot  
```{r}
df = all_leaves_18jun %>%
  separate(tree, c('row', 'tree'))  %>%
  mutate(tree = as.numeric(tree)) %>%
    mutate(row = as.numeric(row))

 df %>%  dplyr::filter(mite_species ==c ("tetranychus", "predatory")) %>%
 arrange(desc(value)) %>%
ggplot(aes(x=row, y=tree, size=value, color=mite_species)) +
    geom_point(alpha=0.5) +
    scale_size(range = c(2, 20), name="Mite count") +
    scale_y_continuous(breaks=seq(0,70,5)) +   
     scale_x_reverse(breaks=seq(20,1)) +
   theme(text = element_text(size=20)) +
    labs(title = "Mite count per leaf, 18/6/2023") +
   theme(panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey90"),  panel.grid.minor.x = element_blank(),  panel.grid.minor.y = element_blank()) 
```


## 7 July 2023
#### as numeric
```{r}
meta_7jul_num <- read.csv("/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/data/7jul_num.csv")
design <- read.csv("/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/data/design.csv")

data_7jul_num <- right_join(design,filter(meta_7jul_num,row !=c("21", "22", "23")), by = "row") %>% 
  # match the cultivar name to each row , and remove rows 21,22,23 (not tested in this date)
  mutate(across(everything(), as.character)) 

 data_7jul_num[data_7jul_num == ""] <- NA
 data_7jul_num[is.na(data_7jul_num)] <- 0

# count adult mites 
#on 5 outer leaves:
out_leaves <- data_7jul_num %>% 
   #dplyr::mutate(across(everything(), as.character)) %>%
  dplyr::select(-contains('egg')) %>%
  dplyr::select(-starts_with('inner')) %>%
  pivot_longer(cols = starts_with('outer')) %>% 
  separate(name, c('leave_position', 'leave_number', 'mite_species')) 

in_leaves <- data_7jul_num %>% 
   #dplyr::mutate(across(everything(), as.character)) %>%
  dplyr::select(-contains('egg')) %>%
  dplyr::select(-starts_with('outer')) %>%
  pivot_longer(cols = starts_with('inner')) %>% 
  separate(name, c('leave_position', 'leave_number', 'mite_species')) 

all_leaves_7jul = bind_rows(in_leaves, out_leaves)

# order the levels 
all_leaves_7jul$mite_species <- factor(all_leaves_7jul$mite_species, level=c("tetranychus", "panonychus", "bryobia", "predatory"))
all_leaves_7jul$treatment <- factor(all_leaves_7jul$treatment, level=c("control", "low", "high"))

all_leaves_7jul %>% 
  group_by(treatment, mite_species) %>%
  dplyr::summarise(mean=mean(value),
            sd=sd(value), .groups = 'drop') %>%
  ggplot(aes(y=mean, x=treatment, fill = treatment)) + 
 geom_bar(position = "dodge",
           stat = "identity") +
        labs(title = "Mite count per leaf") +
  #geom_errorbar(aes(ymin=0, ymax=mean+sd)) +
  facet_wrap(~mite_species, scales ="free") + 
  theme_classic() 
  
# by leave position
 all_leaves_7jul %>% 
  group_by(treatment, mite_species,leave_position) %>%
  dplyr::summarise(mean=mean(value),
            sd=sd(value), .groups = 'drop') %>%
  ggplot(aes(y=mean, x=treatment, fill = leave_position)) + 
 geom_bar(position = "dodge",
           stat = "identity")+
  #    geom_errorbar(position = "dodge",aes(ymin=0, ymax=mean+sd)) +
    labs(title = "Mite count per leaf") +
  facet_wrap(~mite_species, scales ="free") + 
  theme_classic() 

# by apple strain 
 all_leaves_7jul %>% 
  group_by(treatment, mite_species,leave_position,strain) %>%
  dplyr::summarise(mean=mean(value),
            sd=sd(value), .groups = 'drop') %>%
  ggplot(aes(y=mean, x=treatment, fill = strain)) + 
 geom_bar(position = "dodge",
           stat = "identity")+
    # geom_errorbar(position = "dodge",aes(ymin=0, ymax=mean+sd)) +
    labs(title = "Mite count per leaf") +
  facet_wrap(~mite_species, scales ="free") + 
  theme_classic() 

#in boxplot the outliers really bias the plots, making it hard to detect differences
#  all_leaves_7jul %>% 
#  group_by(treatment, mite_species) %>%
#  ggplot(aes(y=value, x=treatment, fill = treatment)) + 
#   geom_boxplot() +
#  facet_wrap(~mite_species) + 
#  theme_classic() 
```
 
##### spatial distribution - bubble plot  
```{r}
df = all_leaves_7jul %>%
  separate(tree, c('row', 'tree'))  %>%
  mutate(tree = as.numeric(tree)) %>%
    mutate(row = as.numeric(row))

df %>%  dplyr::filter(mite_species ==c ("tetranychus", "predatory")) %>%
 arrange(desc(value)) %>%
ggplot(aes(x=row, y=tree, size=value, color=mite_species)) +
    geom_point(alpha=0.5) +
    scale_size(range = c(2, 20), name="Mite count") +
    scale_y_continuous(breaks=seq(0,70,5)) +   
     scale_x_reverse(breaks=seq(20,1)) +
   theme(text = element_text(size=20)) +
    labs(title = "Mite count per leaf") +
   theme(panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey90"),  panel.grid.minor.x = element_blank(),  panel.grid.minor.y = element_blank()) 
```

## collect all 3 monitoring dates
```{r}
# we created 3 files, one per monitoring date:
 #all_leaves_5jun, all_leaves_18jun, all_leaves_7jul

# add column with time from application

moni_1 <- all_leaves_5jun %>% dplyr::select(c("row","treatment","strain","day","tree", "mite_species", "value")) %>%
  mutate(tree = as.character(tree)) %>%
    mutate(row = as.character(row)) 

moni_2 <-all_leaves_18jun %>% dplyr::select(c("row","treatment","strain","day","tree", "mite_species", "value")) %>%    mutate(row = as.character(row)) 

moni_3 <-all_leaves_7jul %>% separate(tree, into= c("row","tree")) %>%
dplyr::select(c("row","treatment","strain","day","tree", "mite_species", "value")) 

data = bind_rows(moni_1, moni_2, moni_3)
  
data %>%
  filter(mite_species != "predatory") %>%
  group_by(treatment, mite_species, day) %>%
  dplyr::summarise(mean=mean(value),
            sd=sd(value), .groups = 'drop') %>% 
  ggplot(aes(day, mean, color = mite_species)) +
   geom_line(size = 1.5) +  
  geom_point() +
#  geom_errorbar(aes(ymin=0, ymax=mean+sd)) +
  facet_wrap(~treatment, nrow = 3) + 
  theme_classic() + 
  labs(y="Mean mite count per leaf") +
  scale_x_continuous(breaks=seq(13,50, 5)) +theme(text=element_text(size=21))

data %>%
  #filter(mite_species != "predatory") %>%
  group_by(treatment, mite_species, day) %>%
  dplyr::summarise(mean=mean(value),
            sd=sd(value), .groups = 'drop') %>% 
  ggplot(aes(day, mean, color = mite_species)) +
   geom_line(size = 1.5) +  
  geom_point() +
#  geom_errorbar(aes(ymin=0, ymax=mean+sd)) +
  facet_wrap(~treatment, nrow = 3) + 
  theme_classic() + 
  labs(y="Mean mite count per leaf") +
  scale_x_continuous(breaks=seq(13,50, 5)) +theme(text=element_text(size=21))


data %>%
  filter(mite_species == "predatory") %>%
  group_by(treatment, mite_species, day) %>%
  dplyr::summarise(mean=mean(value),
            sd=sd(value), .groups = 'drop') %>% 
  ggplot(aes(day, mean, color = mite_species)) +
   geom_line(size = 1.5, color = "#69b3a2") +  
  geom_point(color = "#69b3a2") +
#  geom_errorbar(aes(ymin=0, ymax=mean+sd)) +
  #facet_wrap(~treatment, nrow = 3) + 
  theme_classic() + 
  labs(y="Mean mite count per leaf") +
  scale_x_continuous(breaks=seq(13,50, 5)) +theme(text=element_text(size=21))


data %>%
  filter(mite_species == "predatory") %>%
  group_by(treatment, mite_species, day) %>%
  dplyr::summarise(mean=mean(value),
            sd=sd(value), .groups = 'drop') %>% 
  ggplot(aes(day, mean, color = treatment)) +
   geom_line(size = 1.5) +  
  geom_point() +
#  geom_errorbar(aes(ymin=0, ymax=mean+sd)) +
  #facet_wrap(~treatment, nrow = 3) + 
  theme_classic() + 
  labs(y="Mean mite count per leaf") +
  scale_x_continuous(breaks=seq(13,50, 5)) +theme(text=element_text(size=21))
```







## as order
```{r}
meta_7jul <- read.csv("/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/data/7jul.csv")
design <- read.csv("/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/data/design.csv")

data_7jul <- right_join(design,filter(meta_7jul,row !=c("21", "22", "23")), by = "row") %>% 
  # match the cultivar name to each row , and remove rows 21,22,23 (not tested in this date)
  mutate(across(everything(), as.character)) 

 data_7jul[data_7jul == ""] <- NA
 data_7jul[is.na(data_7jul)] <- 0

 
 #write the arranged data as csv file
#write_csv(x= data_7jul, file="/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/outfiles/data_7jul.csv")

#data <- left_join(design,meta, by = "row") %>% # match the cultivar name to each row
 # filter(row !="23")
 #data <- data %>% mutate(across(-1,  ~ as.numeric(replace(., . == '', 0))))  %>% 
#  mutate(across(everything(), as.character)) %>%
#replace(is.na(.),0)
```

### Adult mites distribution on leaves
#### on outer leaves
```{r}
# count adult mites 
#on 5 outer leaves:
out_leaves <- data_7jul %>% 
   dplyr::mutate(across(everything(), as.character)) %>%
  dplyr::select(-contains('egg')) %>%
  dplyr::select(-starts_with('inner')) %>%
  pivot_longer(cols = starts_with('outer')) %>% 
  separate(name, c('leave_position', 'leave_number', 'mite_species'))  %>% 
  group_by( mite_species, tree, strain, treatment) %>%
  dplyr::count(value, .drop = FALSE) %>%
  mutate(leave_position = "outer_leave")

#on 5 outer leaves:
in_leaves <- data_7jul %>% 
   dplyr::mutate(across(everything(), as.character)) %>%
  dplyr::select(-contains('egg')) %>%
  pivot_longer(cols = contains('inn')) %>% 
  separate(name, c('leave_position', 'leave_number', 'mite_species'))  %>% 
  group_by( mite_species, tree, strain, treatment) %>%
  dplyr::count(value, .drop = FALSE) %>%
  mutate(leave_position = "inner_leave")

#all leaves:
leaves = 
  bind_rows(in_leaves, out_leaves)
```

```{r}
# plot the distribution 
out_leaves$value <- factor(out_leaves$value, levels = c("0","1", "2","3","4","5+","10+", "20+", "30+"))
out_leaves$mite_species <- factor(out_leaves$mite_species, level=c("tetranychus", "panonychus", "bryobia", "predatory"))
out_leaves$treatment <- factor(out_leaves$treatment, level=c("control", "low", "high"))

# scales::show_col(brewer.pal(9, name = "PiYG")) # show the codes of colors for 9 levels

out_leaves %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "mite infestation distribution on *outer leaves*,
comparison between the 4 species") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + 
  facet_wrap(~mite_species) + scale_fill_manual(values=c("#66b032", "#ffeda0","#fed976","#fed24c","#fd8d3c", "#fc4e2a", "#e31a1c","#bd0026","#800026")) 


out_leaves %>% 
  ggplot(aes(fill=value, y=n, x=mite_species)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "mite infestation distribution on *outer leaves*,
comparison between the 4 species") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15,angle = 45, hjust=1),
        axis.title.x = element_blank()) + 
  facet_wrap(~treatment) + scale_fill_manual(values=c("#66b032", "#ffeda0","#fed976","#fed24c","#fd8d3c", "#fc4e2a", "#e31a1c","#bd0026","#800026")) 

out_leaves %>% filter(mite_species == "predatory") %>%
# separate(tree, into= c("row","tree")) %>%
  ggplot(aes(fill=value, y=n, x=mite_species)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "mite infestation distribution on *outer leaves*,
comparison between the 4 species") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank()) + 
  facet_wrap(~row+tree) + scale_fill_manual(values=c("#66b032", "#ffeda0","#fed976","#fed24c","#fd8d3c", "#fc4e2a", "#e31a1c","#bd0026","#800026")) 

out_leaves %>% 
  ggplot(aes(fill=value, y=n, x=mite_species)) + 
geom_boxplot() +  facet_wrap(~treatment) + scale_fill_manual(values=c("#abdda4", "#ffeda0","#fed976","#fed24c","#fd8d3c", "#fc4e2a", "#e31a1c","#bd0026","#800026")) 

out_leaves %>%
  ggplot(aes(fill=value, y=n, x=treatment)) + 
geom_boxplot() + scale_fill_manual(values=c("#abdda4", "#ffeda0","#fed976","#fed24c","#fd8d3c", "#fc4e2a", "#e31a1c","#bd0026","#800026")) +  facet_wrap(~mite_species) 


out_leaves %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "mite infestation distribution on outer leaves
         comparison between the 4 apples cultivars") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + scale_fill_manual(values=c("#66b032", "#FFC300","#FF5733","#C70039")) +
  facet_wrap(~strain)
```

heatmap
```{r}
in_leaves %>% filter(mite_species != "predatory") %>% 
   filter(value !="30+") %>%
  separate(tree, into= c("row","tree")) %>% 
  dplyr::mutate_at(c("tree","n","row"), as.integer) %>% 
ggplot(aes(row, tree, fill= n)) + 
  geom_tile() +scale_fill_distiller(palette = "YlOrRd", trans = "reverse") +
  labs(title = "leaf infestation level per tree, 
*inner leaves*")

 in_leaves %>% filter(mite_species == "predatory") %>% 
   filter(value = c("1","2","3")) %>%
  separate(tree, into= c("row","tree")) %>% 
  dplyr::mutate_at(c("tree","n","row"), as.numeric) %>% 
ggplot(aes(row, tree, fill= n)) + 
  geom_tile() +scale_fill_distiller(palette = "YlOrRd", trans = "reverse") +
  labs(title = "leaf infestation level per tree, 
*inner leaves*")
 
 
out_leaves %>% filter(mite_species == "predatory") %>% 
   filter(value !="0") %>%
  separate(tree, into= c("row","tree")) %>% 
  dplyr::mutate_at(c("tree","n","row"), as.numeric) %>% 
ggplot(aes(row, tree, fill= n)) + 
  geom_tile() +scale_fill_distiller(palette = "YlOrRd", trans = "reverse")+
  labs(title = "leaf infestation level per tree, 
*outer leaves*")


#  PiYG

```

## 18 June 2023
### load data
```{r}
meta_18jun <- read.csv("/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/data/18jun.csv")
design <- read.csv("/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/data/design.csv")

data_18jun <- left_join(design,meta_18jun, by = "row") %>% # match the cultivar name to each row
  filter(row !="23") %>% # remove row 23 (we checked only once)
  mutate(across(everything(), as.character)) 

 data_18jun[data_18jun == ""] <- NA
 data_18jun[is.na(data_18jun)] <- 0

 #write the arranged data as csv file
#write_csv(x= data, file="/Users/nuriteliash/Documents/GitHub/Biological-control-mites-apples/outfiles/data.csv")

#data <- left_join(design,meta, by = "row") %>% # match the cultivar name to each row
 # filter(row !="23")
 #data <- data %>% mutate(across(-1,  ~ as.numeric(replace(., . == '', 0))))  %>% 
#  mutate(across(everything(), as.character)) %>%
#replace(is.na(.),0)
```

### Adult mites distribution on leaves
#### on outer leaves
```{r}
# count adult mites 
#on 5 outer leaves:
out_leaves <- data_18jun %>% 
  dplyr::select(-contains('egg')) %>%
  dplyr::select(-starts_with('inner')) %>%
  pivot_longer(cols = starts_with('outer')) %>% 
  separate(name, c('leave_position', 'leave_number', 'mite_species'))  %>% 
  group_by( mite_species, tree, strain, treatment) %>%
  dplyr::count(value, .drop = FALSE)

# plot the distribution 
out_leaves$value <- factor(out_leaves$value, levels = c("0","1-5","5-10","10+"))

out_leaves %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "mite infestation distribution on outer leaves,
comparison between the 4 species") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + scale_fill_manual(values=c("#66b032", "#FFC300","#FF5733","#C70039")) +
  facet_wrap(~mite_species)

out_leaves %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "mite infestation distribution on outer leaves
         comparison between the 4 apples cultivars") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + scale_fill_manual(values=c("#66b032", "#FFC300","#FF5733","#C70039")) +
  facet_wrap(~strain)
```

#### on Inner leaves
```{r}
# count adult mites 
#on 5 outer leaves:
in_leaves <- data %>% 
  dplyr::select(-contains('egg')) %>%
  dplyr::select(-starts_with('outer')) %>%
  pivot_longer(cols = starts_with('inner')) %>% 
  separate(name, c('leave_position', 'leave_number', 'mite_species'))  %>% 
  group_by( mite_species, tree, strain, treatment) %>%
  dplyr::count(value, .drop = FALSE)

# plot the distribution 
in_leaves$value <- factor(in_leaves$value, levels = c("0","1-5","5-10","10+"))

in_leaves %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "mite infestation distribution on inner leaves
         comparison between the 4 mite species") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + scale_fill_manual(values=c("#66b032", "#FFC300","#FF5733","#C70039")) +
  facet_wrap(~mite_species)

in_leaves %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "mite infestation distribution on inner leaves
         comparison between the 4 apple cultivars") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + scale_fill_manual(values=c("#66b032", "#FFC300","#FF5733","#C70039")) +
  facet_wrap(~strain)
```

### Mites' eggs distribution on leaves
#### on outer leaves
```{r}
# count  mites eggs on 5 outer leaves:
out_leaves_eggs <- data %>% 
  dplyr::select( c(row, treatment, strain, tree),contains('Egg')) %>%
  dplyr::select(-starts_with('inner')) 

# make a list of the columns to keep:
keep <- out_leaves_eggs %>% dplyr::select(starts_with('outer')) %>%
  colnames()
out_leaves_eggs <- out_leaves_eggs %>% 
  mutate_at(keep, as.character) %>%
  pivot_longer(cols = starts_with('outer')) %>% 
  separate(name, c('leave_position', 'leave_number', 'egg_type'))  %>%
  group_by(egg_type, tree, strain, treatment) %>%
  dplyr::count(value, .drop = FALSE)

# plot the distribution 
out_leaves_eggs$value <- factor(out_leaves_eggs$value, levels = c("0","1-5","5-10","10+"))

out_leaves_eggs %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "Egg distribution on outer leaves") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + scale_fill_manual(values=c("#66b032", "#FFC300","#FF5733","#C70039")) +
  facet_wrap(~egg_type)

out_leaves_eggs %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "Egg distribution on outer leaves") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + scale_fill_manual(values=c("#66b032", "#FFC300","#FF5733","#C70039")) +
  facet_wrap(~strain)
```

#### on inner leaves
```{r}
# count  mites eggs  
#on 5 inner leaves:
in_leaves_eggs <- data %>% 
  dplyr::select( c(row, treatment, strain, tree),contains('egg')) %>%
  dplyr::select(-starts_with('outer')) %>%
  pivot_longer(cols = starts_with('inner')) %>% 
  separate(name, c('leave_position', 'leave_number', 'egg_type'))  %>% 
  group_by(egg_type, tree, strain, treatment) %>%
  dplyr::count(value, .drop = FALSE)

# plot the distribution 
in_leaves_eggs$value <- factor(in_leaves_eggs$value, levels = c("0","1-5","5-10","10+"))

in_leaves_eggs %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "Egg distribution on outer leaves") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + scale_fill_manual(values=c("#66b032", "#FFC300","#FF5733","#C70039")) +
  facet_wrap(~egg_type)

in_leaves_eggs %>% 
  ggplot(aes(fill=value, y=n, x=treatment)) + 
    geom_bar(position="fill", stat="identity") +
    xlab("Treatment") + 
    ylab("Outer leaves count per tree") +
    labs(title = "Egg distribution on outer leaves") +
    labs(fill = "Infestation 
     level") + 
  theme_classic() +
  theme(axis.text.x = element_text(size=15),
        axis.title.x = element_blank()) + scale_fill_manual(values=c("#66b032", "#FFC300","#FF5733","#C70039")) +
  facet_wrap(~strain)
```
